---
title: "First try at Mupdog"
author: "David Gerard"
date: "December 14, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

I try out an inefficient version of mupdog for the first time.

# Analysis

```{r}
library(tidyverse)
library(mupdog)
refmat <- as.matrix(read.csv("/home/david/Code/reproduce_genotyping/Output/shirasawa_snps/example_refcounts.csv", row.names = 1))
sizemat <- as.matrix(read.csv("/home/david/Code/reproduce_genotyping/Output/shirasawa_snps/example_readcounts.csv", row.names = 1))
```

Subset since I think this will be slow

```{r}
nind <- nrow(refmat)
nsnps <- 500
refmat_sub <- refmat[1:nind, 1:nsnps]
sizemat_sub <- sizemat[1:nind, 1:nsnps]
```

Try it out!
```{r, cache = TRUE}
mout <- mupdog::mupdog(refmat = refmat_sub, sizemat = sizemat_sub, ploidy = 6, verbose = FALSE, 
                       control = list(obj_tol = 10^-4))
```

# Look at genotyping

The correlation matrix looks a little different than what I expected --- I expected all of the correlations to be about the same.
```{r}
hist(mout$cor_mat[upper.tri(mout$cor_mat)])
```

```{r}
hist(mout$seq)
```

```{r}
hist(log(mout$bias))
```

```{r}
hist(mout$od)
```

Inbreeding doesn't seem to be too important for these data
```{r}
hist(log(mout$inbreeding))
```


The Genotypes all look fantastic.
```{r}
for (index in 1:100) {
  pl <- updog::plot_geno(ocounts = refmat_sub[, index], osize = sizemat_sub[, index], ploidy = 6, 
                         ogeno = mout$map_dosage[, index], bias_val = mout$bias[index], seq_error = mout$seq[index])
  print(pl)
}

```


Some EDA
```{r}
hist(mout$postmean)
hist(mout$inbreeding)
hist(mout$postvar)
mout$seq
mout$bias
mout$od 
```

```{r}
mout$cor_mat
```





```{r}
sessionInfo()
```















